##Assignment 1
###Loading and preprocessing the data
```{r}
library(knitr, XML)
act <- read.csv(file="activity.csv")
```

###What is mean total number of steps taken per day?
```{r,results='asis'}
library(xtable)
means <- aggregate(act[,1], list(act[,2]), mean)
medians <- aggregate(act[,1], list(act[,2]), median)
tab<-cbind(means, medians[2])
colnames(tab) <- c("date", "mean", "median")
tab <- xtable(tab)
```

```{r, results='asis'}
print(tab, type="html")
hist(means[,2], main="Steps taken per day, means", xlab="")
```

###What is the average daily activity pattern?
```{r, results='asis'}
dp <- by(act[,1], act[,3], mean,  simplify=T, na.rm=T)
dp <- as.data.frame(as.table(dp), stringsAsFactors=F)
plot(dp[,1], dp[,2], type="l", xlab="interval", ylab="steps",main="Average daily activity pattern")
```

**`r dp[dp[,2]==max(dp[,2]), 1]`** interval contains the maximum number of steps (`r max(dp[,2])`)

###Imputing missing values

Data contains `r table(is.na(act))[2]` NAs.
To fill empty fields we will use average for interval.

```{r, results='asis'}

act_new <-act
act_new[is.na(act_new),1] <- dp[,2, act_new[,3]==dp[1]]
means_new <- aggregate(act_new[,1], list(act_new[,2]), mean)
medians_new <- aggregate(act_new[,1], list(act_new[,2]), median)
tab<-cbind(means_new, medians_new[2])
colnames(tab) <- c("date", "mean", "median")
tab <- xtable(tab)
```

***Average number of steps per day***

```{r, results='asis'}
print(tab, type="html")
par(mfrow=c(1,2))
hist(means[,2], main="Steps taken per day, means",
     xlab="", ylim=c(1,30))
hist(means_new[,2], main="Steps taken per day, means_new",
     xlab="", ylim=c(1,30))

```

As we can see, imputing missing values changes resulting numbers.

###Are there differences in activity patterns between weekdays and weekends?
```{r}
library(lattice)
Sys.setlocale(category = "LC_TIME", locale = "USA")
act_new$date <- strptime(act_new$date, format="%Y-%m-%d")
act_new$wd <- ifelse(weekdays(act_new[,2])%in%c("Saturday", "Sunday"), "weekend", "weekday")
wdp <- aggregate(act_new[,1], list(act_new[,3], act_new[,4]), mean)
xyplot(wdp[,3] ~  wdp[,1]| wdp[,2], layout=c(1,2), type="l", xlab="interval", ylab="steps", main="Activity patterns for weekdays and weekend")
```

As we can see, activity patterns for weekdays and weekends are different.
Weekand pattern has more distributed activity spikes.